Categorical and gradient aspects of wordlikeness\thanks{Thanks to Gene Buckley, Charles Yang, and audiences at NELS 43 for helpful comments.}

### 2012-11-05 4.tex
Writing on chapter 2 primarily, also added the two "MSes" and new graphs for chapter 2.

\chapter{Structural and accidental phonotatic gaps}
\label{gaps}

Early work in generative phonology notes that neutralizing, surface-true phonological alternations impose constraints both on surface forms and underlying representations (\citealp[283]{Anderson1974}, \citealp[382]{SPE}, \citealp[205f.]{Dell1973}, \citealp[22f.]{SPR}, \citealp{Kisseberth1970b}, \citealp{Kenstowicz1977}: chap.~3, \citealp[28f.]{Stampe1973}, \citealp[410f.]{Stanley1967}).

Two domains are essential to defining

Despite isolated claims to the contrary (e.g., \citealp[297]{Hale1965}; \citealp[212f.]{Postal1968}),

\begin{quote}
It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained. \citep[283]{Anderson1974}
\end{quote}

\section{English syllable contact clusters}

English admits a variety of word-medial consonant clusters spanning syllable boundaries; these are known as \emph{syllable contact clusters} or \emph{interludes}. \citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in some language, echoing a similar proposal by \citet{Haugen1956}.

\begin{quote}
\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
\end{quote}

Many phonotactic generalization can be thought of as restrictions on the inventory of different prosodic constituents defining what is, e.g., a ``well-formed onset''. Since few linguists posit the interlude as a prosodic constituent \citep[though cf.][]{Steriade1999}, further constraints on syllable contact clusters, however, are combinatoric.

@@

@@

@@

The baroque treatment of Latinate forms in \emph{The sound pattern of English} (\citealp{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition to simplify various morphophonological generalizations. In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.

@@

There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}. 
%\citeauthor{Harley2009} proposes that the Latinate prefix is introduced in a low small clause, the same position as the theme, participle, and result adjectival in the Anglo-Saxon examples.

@@

So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters. These constraints are ``static'' in the sense that they have no analogues in the morphophonemics of English. The investigation considers clusters of all sizes, not merely the triconsontal clusters studied by \citeauthor{Pierrehumbert1994}, as she gives no justification for excluding biconsontal or quadraconsonantal clusters.

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 25       & 91         & 22\%       & \multirow{2}{*}{0.106} \\
violating  &  4       & 40         &  9\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{dltab}
\end{table}

\paragraph{Coronal codas} \citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}. Table \ref{cctab} shows that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}). Restricting attention to triconsonantal clusters does not produce a significant lexical effect, either ($p = 0.129$).

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 56       & 304        & 15\%       & \multirow{2}{*}{0.430} \\
violating  & 37       & 243        & 13\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{cctab}
\end{table}

\noindent

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 47       & 512        &  8\%       & \multirow{2}{*}{0.250} \\
violating  &  0       &  25        &  0\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{abatab}
\end{table}

@@

\paragraph{Obstruent voice assimilation}

Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), which take the voicing specification of a preceding obstruent.\footnote{Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 35       & 329        & 10\%       & \multirow{2}{*}{0.002} \\
violating  & 11       & 305        &  3\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{ovatab}
\end{table}

@@

\textsc{Nasal Place Assimilation} (\emph{SPE}:85) accounts for [n $\sim$ ŋ] allophony and Latinate prefix allomorphy, such as that of the ``negative'' refix \emph{im-}/\emph{in-}.\footnote{\citealt{Borowsky1986}:FIXME and \citealt{Jensen2000} propose additional restrictions on \textsc{Nasal Place Assimilation}.}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.37\e{-07}} \\
violating  & 11       & 22         & 33\%       \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{npatab}
\end{table}

\paragraph{Degemination} The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealp[102]{Borowsky1986}). \textsc{Degemination} is formalized here as a rule deleting the first of two identical segments, ignoring voice.

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
violating  & 0        & 104        & 0\%        \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{degemtab}
\end{table}

\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters.

@@

A soft-margin support vector machine \citep{Cortes1995} with a linear kerne
l is used to convert from the per-cluster scores to attestation predictions
, by finding a single cutoff which optimally separates attested and unattes
ted clusters according to their model scores. This is in no way intended to
 be a cognitively plausible model for learning phonotactics; it simply repr
esents an upper bound on the performance of these models. The results for a
ll four models are summarized in Table \ref{cmresults}.

\begin{table} 
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
DC and EF           & 0.861    & 0.868     & 0.988  & 0.924 \\
\bottomrule
\end{tabular}
\caption{FIXME}
\label{cmresults}
\end{table}

@@

\subsubsection{Simulating medial clusters}

@@

\begin{figure}
\centering
\includegraphics{sim.pdf}
\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line). Both simulated and observed frequencies have been smoothed using the $Z_r$ transform.}
\label{sim}
\end{figure}

@@

@@

%however, the \emph{communis opinio} still is that that not all phonotactics---that is, language-specific sound sequence co-occurrence restrictions---can be derived in this fashion, and phonotactic knowledge is independent of the phonology.
%
%%FIXME
%% \citet[255]{Anderson1974}, discussing this fact, writes that \citeauthor{Haugen} ``gives an argument that shows the necessity of including syllables as units in underlying representations''.
%
%%FIXME
%(for dissenting views, see \citealt{Blevins2003} and \citealt{Steriade1999})
%
%Linguists have long recognized underlying representations as privileged domains for stating phonotactic generalizations (e.g., \citealt{Bloomfield1930,Jakobson1932,Pike1947b}). 
%
%More recently, a similar role for the syllable has been recognized 
%(e.g., \citet{Haugen1956}, \citet{Hooper1973}, \citealt{Kahn1976})
%
%
%\begin{quote}
%It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained. \citep[283]{Anderson1974}
%\end{quote}
%
%
%\begin{quote}
%It would be a result of the greatest interest if it were to turn out that every intramorphemic regularity was necessarily reflected in an alternation, and thus in a rule; but this position cannot seriously be maintained. \citep[283]{Anderson1974}
%\end{quote}
%
%English word-medial consonant clusters have been used, by \citet{Pierrehumbert1994} and others, to argue for the necessity of an independent phonotactics module and illustrate proposals for its architecture. This chapter argues, however, that the only structural constraints on the concatenation of codas and onsets to form word-medial clusters are derivative of well-documented phonological alternations in English.
%
%\section{Background}
%
%English admits a variety of word-medial consonant clusters spanning syllable boundaries, known as \emph{syllable contact clusters} or \emph{interludes}. These clusters can be as short as two consonants (e.g., \emph{a}[n.t]\emph{ics}; singleton intervocalic consonants are ignored here) or as long as four (e.g., \emph{mi}[n.str]\emph{el}). \citeauthor{Pierrehumbert1994} presents the following as a null hypothesis for which clusters are admissible in any given language \citep[cf.][]{Haugen1956}.
%
%\begin{quote}
%\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
%\end{quote}
%
%This chapter is concerned with the nature of these ``provisos'' as reflected by which of the ``possible'' clusters, so defined, are attested. While many phonotactic generalizations can be thought of as inventory restrictions imposed by prosodic constituents, additional filters on the set of possible clusters are essentially combinatoric in nature and thus distinct, there is no prosodic constituent consisting of the coda of one consonant-final syllable and the onset of a following consonant-initial syllable \citep[though see][]{Steriade1999}.
%
%\section{Evaluation}
%
%\citeauthor{Pierrehumbert1994} argues that the inventory of clusters in English shows the effect of a small number of static phonotactic constraints. The remainder of this chapter provides an exhaustive quantitative evaluation of this hypothesis. The most important constraints on medial clusters are those derived from phonological rules which apply across syllable boundaries; there is no remaining role for static constraints. The sparse nature of the lexicon plays a secondary role, giving rise to apparent phonotactic gaps even in the absence of static phonotactic constraints.
%
%\subsection{Corpus construction}
%
%A list of medial clusters is gathered by applying a number of filters and transformations to an English lexical database. The resulting list of clusters is reproduced in Appendix \ref{appendixC} and in electronic form at the author's website.
%
%\subsubsection{Simplex words}
%
%An analysis of lexical constraints requires a corpus of lexical representations. So as to remain agnostic about particulars of morphological theory, the goal here is to identify words which bear no overt inflection, and which do not admit any plausible further decomposition, henceforth ``simplex words''. For the structuralists, a procedure for parsing words into morphs, just the thing for locating simplex words, held the status of ``philosopher's stone'', a sort of near-mythical objective of the endeavor, but actual attempts to develop such a procedure \citep[e.g.,][]{Harris1955,Nida1948} are now recognized as anything but theory-neutral. Yet, there is no alternative to adopting some computational procedure or human-coded database as the ``gold standard'' for a study of this type. \citet{Pierrehumbert1994} analyzes a list of entries in the Collins English dictionary by hand, using her intuitions to filter out ``complex'' words as well as those which are not ``reasonably familiar''. Unfortunately, the resulting list was neither published nor circulated, and the possibility of replicating \citeauthor{Pierrehumbert1994}'s' sensations of morphological complexity is remote. Following prior studies of syllable contact clusters by \citet[ chapter 3]{Hammond1999a} and \citet[ chapter 8]{Duanmu2009}, a list of simplex words was derived from the ``lemmas'' (that is, uninflected forms) in the English portion of CELEX \citep{CELEX}, a database of morphological, phonological, and syntactic annotations based on the COBUILD corpus, a resource that was also used to construct the Collins English dictionary employed by \citeauthor{Pierrehumbert1994}.
%
%Two further filters were applied to this list. First, any lemma coded as an unassimilated loanword is excluded. Secondly, all lemmas that have a ``morphological status'' other than ``monomorpheme'' are excluded. This is crucial because CELEX makes no guarantee that lemmas are simplex; indeed, many lemmas are products of derivation (e.g., \emph{abusive}, from \emph{abuse}). These stringent criteria have a profound effect on the makeup of the data. For instance, nearly all exceptions to \textsc{Obstruent Voice Assimilation} (see \S\ref{ova} below) cited by \citet[74]{Hammond1999a} are excluded as unassimilated loanwords (e.g., \emph{vodka}, \emph{smorgasbord}) or as morphologically complex (e.g., \emph{jurisdiction}, \emph{madcap}, \emph{tadpole}, \emph{scapegoat}, \emph{magpie}). 
%
%``Neo-classical compounds'', words that appear to consist of a Latinate prefix and a bound stem, like \emph{inspect} or \emph{excrete}, have an ambiguous and long-debated morphological status. \citeauthor{Pierrehumbert1994} attempts to distinguish between words like \emph{complete}, \emph{extreme}, \emph{inspect}, \emph{obtuse}, which she assumes to be simplex, and \emph{excrete} and \emph{transparent}, which she considers complex. However, CELEX codes these words as complex and they are thus excluded. This is consistent with converging evidence that speakers decompose these Latinate forms.

%Decomposition is also supported by lexical decision results. Among nonce words, those that can be exhaustively decomposed into prefix and bound stem, like *\emph{de}-\emph{juvenate}, are processed slower that contain a licit morph but cannot be exhaustively decomposed, like *\emph{de-pertoire}, and these in turn are processed slower than those which admit no such decomposition \citep{Taft1975}.\footnote{The same contrasts are present in many other domains where decomposition into root and affix(es) is less controversial, for instance, inflected verbs in Italian \citep{Caramazza1988}.} \citet{Emmorey1989} reports facilitative auditory priming between pairs of words which share a bound stem, like \emph{permit}-\emph{submit} \citep[though see][]{Marslen-Wilson1994}.\footnote{It may be the case that full decomposition is only one of the two methods for the processing of complex words, the other being whole word looking, either running in serial \citep{Caramazza1988} or in parallel \citep{Baayen1997b}. Whatever evidence that is amassed in support of a whole word lookup procedure, however, does not undermine the considerable evidence for decomposition.}
%

>>>>

\label{gaps}

%\footnote{While there is a contentious debate as to whether devoicing in German is completely neutralizing \citep[e.g.,][]{Fourakis1984} or not \citep[e.g.,][]{Port1985}, it is not strictly relevant to the discussion at hand.}

Since at least \citet{Pike1947b}, linguists have posited language-specific constraints on the contents of underlying representations.
These \emph{morpheme structure constraints} were at one time thought to fully account for speakers' knowledge of possible and impossible words (e.g., \citealt{Chomsky1965}, \citeyear[382]{SPE}, \citealt[22f.]{SPR}, \citeyear{Halle1962}, \citealt{Stanley1967}). 
\citet{Shibatani1973}, however, argues that morpheme structure constraints cannot account for all wordlikeness constraints.
In German, for instance, final obstruents devoice, and as a result, there are pairs such  [ɡʀaːt]-[ɡʀaːtə] `ridge(s)' and [ɡʀaːt]-[ɡʀaːdə] `degree(s)' differing only in the plural shape of the root.
Since the voicing of final obstruents is contrastive, the restriction on the voicing of obstruents cannot be a morpheme structure constraint. 
Yet, \citeauthor{Shibatani1973} claims that native speakers reject nonce words ending in final voiced obstruents ``\emph{on the ground that they end in voiced obstruents}'' (95). 
Thus, not all constrasts in possible wordhood can be stated as morpheme structure constraints.

Whereas \citeauthor{Shibatani1973} maintains that morpheme structure constraints are insufficient to account for speakers' knowledge of possible wordhood, others argue that morpheme structure constraints are also unnecessary. \citet[297]{Hale1965}, \citet{Kisseberth1970b}, and \citet[212f.]{Postal1968} observe that the structural descriptions of phonological processes often are often reflected in the lexical redundancies in the same language. In Russian, there are alternations implicating a process of obstruent voice assimilation, and tautomorphemic obstruent clusters have uniform voicing \citep[283]{A74}. \citet[205f.]{Dell1973} and \citet[28f.]{Stampe1973} argue that morpheme structure constraints are otiose, as phonological rules triggering alternations impose restrictions on the contents of URs, an effect known as \emph{Stampean occultation}.

There are some apparent constraints on URs which are not reflected in the system of alternations; these generalizations seem to require reference to the units of prosody (\citealt{Hooper1973}, \citealt{Kahn1976}; see \citealt{Blevins2003} and \citealt{Steriade1999} for possible alternatives). Since word-level prosodic structure is generally thought to be largely redundant and thus absent from underlying representations, these cannot be equated with morpheme structure constraints either. \citeauthor{Haugen1956} notes, for instance, that word-medial consonant cluster inventory is not arbitrary in prosodic terms.

\begin{example}[\textsc{Medial Cluster Law} (after \citealt{Haugen1956})]
\label{mcl}
A word-medial consonant cluster consists of at most a well-formed medial coda followed by a well-formed medial onset
\end{example}

Reviewing the inventory of word-medial consonant clusters in English, 
\citet{Pierrehumbert1994} notes that the vast majority of ``possible clusters'', those which can be parsed into a coda and onset, are unattested. \citeauthor{Pierrehumbert1994} argues that the cluster inventory is further filtered by what are essentially morpheme structure constraints: static co-occurrence restrictions with no support from either prosody or the system of alternations. This study will argue, however, that the only structural constraints on word-medial clusters in English are the \textsc{Medial Cluster Law} and Stampean occultation. Consequently, this domain provides no evidence for the existence of morpheme structure constraints.

\section{English syllable contact clusters \citep{Pierrehumbert1994}}

English admits a variety of word-medial consonant clusters spanning syllable boundaries which are also known as \emph{syllable contact clusters} or \emph{interludes}. These clusters can be as short as two consonants, as in \emph{a}[n.t]\emph{ics}, or as long as four, as in \emph{mi}[n.str]\emph{el}. 
There are three

\subsection{The role of Stampean occultation}

Finally, \citeauthor{Pierrehumbert1994} does not fully consider the possible effects of Stampean occultation, as some of the static constraints she proposes may be related to English morphophonemics. For instance, she writes that ``nasal-stop sequences agree in labiality'' (175), and attributes this the status of a static constraint. However, this generalization is just a narrower form of the one derived by \textsc{Nasal Place Assimilation} (see \S\ref{npa}), and which requires no further explanation. Other derived constraints go unnoticed: for instance, she does not discuss the highly reliable tendency of obstruent-obstruent clusters to agree in voicing (see \S\ref{ova}). The evaluation below considers the role of both types of constraints on the English lexicon on an equal footing.

\subsection{The role of sparsity}

\citet{Pierrehumbert1994} infers static constraints from near-exceptionless gaps in the lexicon, but little effort is made to show that the patterns of lexical underrepresentation from which the static constraints are inferred are not due to chance. This admits the possibility that the gaps in the consonant cluster inventory are accidental rather than structural in nature \citep{Fischer-Jorgensen1952,Vogt1954}. This is made all the more likely given the tendency of segment and cluster frequency distributions to be highly skewed \citep[e.g.,][]{Pande2010,Sigurd1968,Tambovtsev2007,Weiss1961}. In Figure \ref{cao}, the rank and type frequency (i.e., frequency in the lexicon) frequency of medial codas and onsets in the lexical sample described below are graphed in log-log space. Both codas and onsets show a linear relationship between log rank and log frequency characteristic of Zipfian distributions (see Appendix \ref{zr}). As a result, an enormous lexical sample would be needed to realize clusters consistent with the \textsc{Medial Cluster Law} even if there were no constraints on the combination of medial codas and onsets in English.

\begin{figure}
\centering
\includegraphics{co.pdf}
\caption{Medial coda and onset type frequencies in the lexical sample show a Zipfian distribution; frequencies have been smoothed using the $Z_r$ transform (see Appendix \ref{zr}).}
\label{cao}
\end{figure}

\subsection{The role of morphological segmentation}

Many components of \citeauthor{Pierrehumbert1994}'s study cannot be directly replicated. \citeauthor{Pierrehumbert1994} focuses her attention on those words she judges to be ``morphologically simple'' and ``reasonably familiar''. Since this list has neither been published nor circulated, any replication will have to be indirect. Further, it is not clear that the judgements (as well as cognitive limitations) of concerned parties should be granted evidential status in the first place, given the potential for implicit bias. A sequence which is rare in simplex words is sometimes said to mark a juncture even in the absense of other evidence for segmentation; for instance, \citet[546]{Rice2009d} analyses many words in Slave as compounds simply because they contain consonant clusters that rarely occur morph-internally. There is some evidence that infants \citep{Mattys2001b} and adults \citep{Brown1956,Hay2004a,McQueen1998b,Norris1997} use this heuristic to segment fluent speech in experimental settings. Applied indiscriminately, though, such a heuristic has the possibility to trivializes both morphological segmentation and phonotactic generalization.

@@

\citeauthor{Pierrehumbert1994} focuses only on triconsonantal clusters, and no reason is given for the exclusion of other sizes of clusters.

@@

%The same practice is sometimes applied to preserve phonological generalizations; for instance, otherwise unmotivated morphological boundaries are used in \emph{SPE} to simplify principles of stress assignment. 
Such reasoning is found in \citeauthor{Pierrehumbert1994}'s discussion of the word \emph{a}[nt.l]\emph{er}, however: this contains a cluster which violates her constraint on coda coronal obstruents, and therefore must be a complex word, she argues (see \citealt[164]{Borowsky1989} for additional exceptions to this generalization).

@@

The treatment of Latinate forms in \emph{The sound pattern of English} (\citealt{SPE}, henceforth \emph{SPE}) assumes a prefix/bound stem decomposition. In a similar vein, \citet[11f.]{Aronoff1976} observes that Latinate forms which share the same bound stem also share irregular allomorphs of that stem under derivation.

@@

There is also a syntactic interaction between Latinate prefixes on verbs and the makeup of the verbal complement. \citet{Harley2009} observes that Latinate verbs do not generally participate in ditransitive, verb participle, or adjectival resultative constructions, all acceptable with similar Anglo-Saxon verbs \citep[see also][]{Gropen1989,Coppock2008}.

@@

So as to account for the 81\% of ``possible'' but unattested clusters, \citet{Pierrehumbert1994} proposes three constraints on English medial clusters. These constraints are ``static'' in the sense that they have no analogues in the morphophonemics of English.

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 25       & 91         & 22\%       & \multirow{2}{*}{0.106} \\
violating  &  4       & 40         &  9\%       \\
\bottomrule
\end{tabular}
\caption{Dorsal-labial cluster attestation in the lexical sample}
\label{dltab}
\end{table}

\paragraph{Coronal obstruent codas} \citet[175]{Pierrehumbert1994} claims that ``clusters with a coronal obstruent in the coda do not occur'', while noting exceptions like \emph{a}[nt.l]\emph{er}, \emph{ke}[s.tr]\emph{el} and \emph{oi}[nt.m]\emph{ent}. Table \ref{cctab} reveals that coda coronal obstruent clusters are not significantly less likely to occur than non-coronal obstruent clusters (e.g., \emph{re}[p.t]\emph{ile}). Restricting attention to triconsonantal clusters does not produce a significant lexical effect, either ($p = 0.129$).

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 56       & 304        & 15\%       & \multirow{2}{*}{0.430} \\
violating  & 37       & 243        & 13\%       \\
\bottomrule
\end{tabular}
\caption{Coda coronal obstruent cluster attestation in the lexical sample}
\label{cctab}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 47       & 512        &  8\%       & \multirow{2}{*}{0.250} \\
violating  &  0       &  25        &  0\%       \\
\bottomrule
\end{tabular}
\caption{ABA cluster attestation in the lexical sample}
\label{abatab}
\end{table}

@@

\paragraph{Obstruent voice assimilation}
\label{ova}

Voice assimilation alternations are evidenced by the non-syllabic allomorphs of the regular past (e.g., \emph{nap}[t] $\sim$ \emph{nab}[d]) and noun plural (e.g., \emph{lap}[s] $\sim$ \emph{lab}[z]), which take the voicing specification of a preceding obstruent.\footnote{Underlying /-d, -z/ are assumed here (e.g., \citealt{Anderson1973a}, \citealt[284f.]{Bakovic2005b}, \citealt{Basboll1972}, \citealt[210]{SPE}, \citealt[282]{Hockett1958}, \citealt[102]{Pinker1988}, \citealt{Shibatani1972}); alternative analyses are put forth by \citet[210f.]{LANGUAGE}, \citet[135]{Borowsky1986}, \citet{Hoard1971}, \citet{Kiparsky1985}, \citet{Lightner1970}, \citet{Luelsdorff1969}, \citet{Miner1975}, \citet[426]{Nida1948}, and \citet{Zwicky1975}.}
%Bizarely, \citet[208]{Wetzels2001}, cite English as an example of a language without ``general devoicing or assimilatory effects''.

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 35       & 329        & 10\%       & \multirow{2}{*}{0.002} \\
violating  & 11       & 305        &  3\%       \\
\bottomrule
\end{tabular}
\caption{Obstruent voice assimilation cluster attestation in the lexical sample}
\label{ovatab}
\end{table}

@@

\textsc{Nasal Place Assimilation} (e.g., \citealt[149]{Borowsky1986}, \emph{SPE}:85, \citealt[62]{Halle1985a}) accounts for [n $\sim$ ŋ] allophony as well as Latinate prefix allomorphy.

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 31       & 2          & 94\%       & \multirow{2}{*}{3.37\e{-07}} \\
violating  & 11       & 22         & 33\%       \\
\bottomrule
\end{tabular}
\caption{Nasal place assimilation cluster attestation in the lexical sample}
\label{npatab}
\end{table}

\paragraph{Degemination} The final alternation found in English medial clusters is the simplification of geminates which is characteristic of ``level I'' morphology, and is found in the irregular /-t/ past tense (e.g., \emph{bend}/\emph{ben}[t], \emph{build}/\emph{buil}[t]) and in \emph{-ly} deadjectival derivatives (e.g., \emph{norma}[l]\emph{y}, cf. \emph{calm}[l]\emph{y}), and Latinate prefix allomorphy (\emph{SPE}:148; \citealt[102]{Borowsky1986}). \textsc{Degemination} is formalized here as a rule deleting the first of two identical segments, ignoring voice.

@@

\begin{table}
\centering
\begin{tabular}{l rrrr}
\toprule
           & attested & unattested & saturation & $p$-value \\
\midrule
conforming & 173      & 643        & 21\%       & \multirow{2}{*}{1.2\e{-10}} \\
violating  & 0        & 104        & 0\%        \\
\bottomrule
\end{tabular}
\caption{Degemination in the lexical sample}
\label{degemtab}
\end{table}

\paragraph{Summary} All three of \emph{SPE} rules targeting medial clusters have a robust effect in constraining the inventory of possible word-medial syllable clusters; possible clusters which are surface exceptions to these three rules are much less likely to be attested than those which conform to them.

@@

A soft-margin support vector machine \citep{Cortes1995} with a linear kernel is used to convert from the per-cluster scores to attestation predictions
, by finding a single cutoff which optimally separates attested and unattes
ted clusters according to their model scores.

It is possible to maximize recall at the expense of precision, by predicting all clusters to be unattested, or to increase precision at the expense of recall, by predicting non-attestation for a great number of clusters. $F_1$, the harmonic mean of these two measures, is a balanced measure of this tradeoff.

This is not put forth as a cognitively plausible model for learning of phonotactics; it simply represents an upper bound for classifying possible clusters.

The results for are summarized in Table \ref{cmresults}.

\paragraph{Null baseline} Since the majority of possible clusters are unattested, a null baseline which predicts all clusters to be unattested achieves classifies 81\% of clusters correctly, achieving perfect recall at the price of precision.

\paragraph{Expected frequency} \citet{Pierrehumbert1994} proposes that the well-formedness of a syllable contact clusters is proportional to the independent probability of the coda and onset that make it up. \citeauthor{Pierrehumbert1994} reports that this is the best predictor of which complex clusters occur, and while it is a significant improvement on the baseline (sign test $p = 4.5$\e{-05}), it is not distinguished in performance from the other models under consideration.

\paragraph{Derived constraints} The three derived generalizations define a simple classifier, predicting attestation for all clusters which do not violate these three generalizations. This is a significant improvement on the null baseline (sign test $p = 4.0$\e{-09}), preserving high recall while improving precision. The predictions of the derived constraints classifier can also be combined with information on expected frequency. As expected, this boosts precision, since FIXME. This combination produces a model with the highest accuracy and $F_1$.

\paragraph{Maximum entropy phonotactics} \citet{Hayes2008a} present a model which uses the principle of maximum entropy to weigh a large number of competing phonotactic constraints. Since this model has many experimenter-determined parameters, a direct replication of the experiments reported by \citealt{Hayes2008a} is attempted; their software, feature specification, and model settings are all used. Training in this model is inherently stochastic, producing slightly different constraints and weights on each run, so the least accurate of ten independent runs is used for evaluation; this is also the practice of \citeauthor{Hayes2008a}. This model performs approximately as well as other classifiers, but has the poorest recall; this suggests that the constraints learned are overly narrow and fail to filter out many impossible clusters. FIXME

\paragraph{Summary} FIXME

\begin{table} 
\centering
\begin{tabular}{l | rrrr}
\toprule
                    & accuracy & precision & recall & $F_1$ \\
\midrule
Baseline            & 0.812    & 0.812     & 1.000  & 0.896 \\
Expected frequency  & 0.836    & 0.845     & 0.984  & 0.909 \\
Derived constraints & 0.838    & 0.840     & 0.997  & 0.912 \\
DC \& EF            & 0.861    & 0.868     & 0.988  & 0.924 \\
\citet{Hayes2008a}  & 0.835    & 0.942     & 0.854  & 0.896 \\
\bottomrule
\end{tabular}
\caption{Results of the cluster classification task; a combination of derived constraints and expected frequency (``DC \& EF'') produces the highest accuracy and $F_1$.}
\label{cmresults}
\end{table}

@@

\subsubsection{Simulating medial clusters}
\label{simulate}

@@

\begin{figure}
\centering
\includegraphics{sim.pdf}
\caption{A simulated cluster inventory, generated by random sampling and applying phonological neutralizations closely matches the observed cluster frequencies (represented by the smoothing line). Simulated and observed frequencies have been smoothed using the $Z_r$ transform.}
\label{sim}
\end{figure}

@@

In summary, the sparse cluster inventory, previously taken as evidence for static constraints on syllable contact clusters, is expected even if the only constraints on possible clusters are those which are phonologically derived.

@@

None of the foregoing results have revealed any evidence for structural constraints on the English syllable contact cluster except those derived from phonological alternations. What then is to be said of the 366 clusters which do not violate a derived constraint, but which are yet unattested, like [b.z] or [z.n]? Since these gaps appear to be quite arbitrary from a phonological perspective---they cannot be reliably discovered either by linguist or computer---it seems reasonable to identify these clusters as accidental gaps, resulting from the sparsity of the sample and of the English lexicon.

%It is quite apparent that root-internal hetero-voiced obstruent clusters like [s.b] or [z.t] are also rare compared to clusters which are uniformly voiced or uniformly voiceless, as predicted the assimilation rule. Yet, \citet{Pierrehumbert1994} does not mention voice assimilation in her study of English syllable contact clusters. \citet[][74f.]{Hammond1999a} cites words like \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}, which contain hetero-voiced obstruent clusters, as evidence that the process does not apply root-internally, though few of his examples are in fact simplex according to CELEX.\footnote{The \textsc{Revised Alternation Condition} (RAC) proposed by \citeauthor{Kiparsky1973a} (\citeyear{Kiparsky1973a}:163, \citeyear{Kiparsky1982a}:152) blocks the application of obligatory neutralization processes like \textsc{Obstruent Voice Assimilation} in root-internal (i.e., non-derived) environments. Simply because there are exceptions in non-derived environments, \textsc{Obstruent Voice Assimilation} is always consistent with the RAC whether or not it actually applies in non-derived environments. This is due to the ``obligatory'' condition of the RAC. If the rule applies in non-derived environments, then it has lexical exceptions (e.g., \emph{a}[s.b]\emph{estos}) and is not obligatory and thus free from the RAC. On the other hand, if the rule is subject to the RAC, it only applies in derived environments, in which it is obligatory.}
% 
%The question is: does \textsc{Obstruent Voice Assimilation} contribute a meaningful characterization the English lexicon? More specifically, are hetero-voiced clusters underrepresented in a way that is unlikely under the hypothesis of free combination? To answer this question, the 720 clusters in which both the final coda consonant and initial onset consonant are obstruent are sorted into four bins, according to whether they are attested or not, and whether or not the conform to the derived constraint---both obstruents are either voiced or voiceless, like in \emph{hu}[z.b]\emph{and} or \emph{rha}[p.s]\emph{osdy}, respectively---or violate it, like the aforementioned \emph{a}[b.s]\emph{inth} and \emph{a}[s.b]\emph{estos}.
